Revisiting~the~Classics: Online~RL~in~the~Programmable~Dataplane

### 2021-05-07 paper/paper.tex
Architecture figures.

\newcommand{\approach}{On Path Learning}
\newcommand{\approachshort}{OPaL}
\newcommand{\coopfw}{parallel}
\newcommand{\Coopfw}{Parallel}
\newcommand{\indfw}{single}
\newcommand{\Indfw}{Single}

@@

\begin{figure}
	\centering
	\resizebox{1.2\linewidth}{!}{\begin{tikzpicture}
			\node (sw2) {};

			\node[fill=white!80!uofgcobalt, draw=black, minimum height=5.5cm, minimum width=2.9cm, below right= 0.1 of sw2.north west](p4-master){};
			\node[below right, inner sep=2pt] at (p4-master.north west) {\small P4 Datapath};
			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.8] (p4-rules) at (p4-master.north) {\begin{varwidth}{2.5 cm}\scriptsize \centering Flow Control Rules\end{varwidth}};
			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n1) at (p4-rules.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Extern-based Action\end{varwidth}};
			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n2) at (n1.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering ...\end{varwidth}};
			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (oswlc) at (n2.south) {\begin{varwidth}{1.5 cm}\scriptsize \centering Control Packet Match\end{varwidth}};
			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n3) at (oswlc.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering ...\end{varwidth}};
			\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, below=0.1] (n4) at (n3.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Network Telemetry\end{varwidth}};

			\node (pkt-text) at ($(p4-master.north) + (0.0,0.6)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Packet In\end{varwidth}};

			\node (pkt-text-out) at ($(p4-master.south) + (0.0,-0.6)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Packet Out\end{varwidth}};

			\node[fill=white!80!uofgsunshine, draw=black, minimum height=5cm, minimum width=5cm](rli) at ($(p4-master.east) + (5, -1)$){};
			\node[below right, inner sep=2pt] at (rli.north west) (rli-text) {\small RL Island};

			\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-out) at ($(rli-text.south) + (0.5,-0.5)$){};
			\node[below right, inner sep=2pt] at (wq-out.north west) {\small EMEM};
			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-in) at (wq-out.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Input Queue\end{varwidth}};

			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (loop) at ($(wq-out.east) + (1.5,-0.5)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Control ME\end{varwidth}};

			\node[draw=black, dotted, minimum height=1cm, minimum width=2cm](wq-act) at ($(loop.north) + (0.0,1.5)$){};
			\node[below right, inner sep=2pt] at (wq-act.north west) {\small EMEM};
			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (wq-actin) at (wq-act.south) {\begin{varwidth}{2.5 cm}\scriptsize \centering Action Queue\end{varwidth}};

			\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-cls-out) at ($(loop.south) + (0,-1.5)$){};
			\node[below right, inner sep=2pt] at (policy-cls-out.north west) {\small CLS};
			\node[above right, inner sep=2pt] at (policy-cls-out.north west) {\small Policy};
			\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-ctm-out) at ($(policy-cls-out.south) + (0,-0.25)$){};
			\node[below right, inner sep=2pt] at (policy-ctm-out.north west) {\small CTM};
			\node[draw=black, dotted, minimum height=0.5cm, minimum width=2cm](policy-emem-out) at ($(policy-ctm-out.south) + (0,-0.25)$){};
			\node[below right, inner sep=2pt] at (policy-emem-out.north west) {\small EMEM};

			\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t1) at (policy-cls-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T1\end{varwidth}};
			\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t2) at (policy-ctm-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T2\end{varwidth}};
			\node[fill=white!90!uofgsunshine, draw, rectangle, above=0.1, below left= 0.05] (t3) at (policy-emem-out.north east) {\begin{varwidth}{2.5 cm}\scriptsize \centering T3\end{varwidth}};

			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-1) at ($(wq-out.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};

			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-2) at ($(po-me-1.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};

			\node[fill=white!90!uofgsunshine, draw, rectangle, rounded corners=0.05cm, above=0.1] (po-me-3) at ($(po-me-2.south) + (0,-1.0)$) {\begin{varwidth}{2.5 cm}\scriptsize \centering Policy ME\end{varwidth}};

			\draw[->] (oswlc.east)  to node[midway, above, sloped] {\tiny Control, Requests} (wq-in.west);
			\draw[->] (n4.east)  to node[midway, above, sloped] {\tiny State} (wq-in.west);
			\draw[<->, dotted] (wq-in.east)  to (loop.west);
			\draw[->, dotted] (loop.south) to (policy-cls-out);

			\draw[->] (loop.north)  to node[midway, above, sloped] {\tiny Actions} (wq-actin.south);

			\draw[->] (wq-actin.west)  to node[midway, above, sloped] {\tiny Actions} (n1.east);

			\draw[<->, dotted] (po-me-1.east) to (policy-cls-out.west);
			\draw[<->, dotted] (po-me-2.east) to (policy-ctm-out.west);
			\draw[<->, dotted] (po-me-3.east) to (policy-emem-out.west);

			\draw[->] (pkt-text)  to  (p4-master.north);
			\draw[->] (p4-master.south)  to  (pkt-text-out);

			\draw[->] (loop) to node[midway, above, sloped] {\tiny NN} (po-me-1);
			\draw[->] (po-me-1) to node[midway, above, sloped] {\tiny NN} (po-me-2);
			\draw[->] (po-me-2) to node[midway, above, sloped] {\tiny NN} (po-me-3);
			\draw[->] (po-me-3) to [out=-45,in=0,looseness=2] node[midway, above, sloped] {\tiny NN} (loop);
	\end{tikzpicture}}
	\caption{Architecture for an \approachshort{} agent on Netronome hardware, using 4 cores for parallel policy processing. (FIXME: redo in illustrator? Show CTXes, PISA, \Coopfw{} vs \Indfw{})\label{fig:netro-arch}}
\end{figure}

@@

\begin{figure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[keepaspectratio, width=\linewidth]{figures/single-core-arch-sketch}
		\caption{\Indfw{} (offline throughput-optimal).\label{fig:single-and-parallel:single}}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[keepaspectratio, width=\linewidth]{figures/parallel-arch-sketch}
		\caption{\Coopfw{} (online-optimal).\label{fig:single-and-parallel:parallel}}
	\end{subfigure}
	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to latency/throughput needs. ?? MAKE ME FOR REAL \label{fig:single-and-parallel}}
\end{figure}

@@

@@

\begin{table}
\caption{NFP memory use due to \approachshort{} using \numlist{1;4} MEs (\SI{32}{\bit}). CLS and CTM regions are shared between all programs on the same island (placing our RL agent on i5), while EMEM and IMEM regions are shared between all NFP programs on a NIC.\label{tab:resources} ??NEEDS REWORKED}
\resizebox{\linewidth}{!}{
\begin{tabular}{@{}cSSSSSSSSSS@{}}
	\toprule Firmware & \multicolumn{2}{c}{EMEM} & \multicolumn{2}{c}{EMEM Cache} & \multicolumn{2}{c}{IMEM} & \multicolumn{2}{c}{i5.CLS} & \multicolumn{2}{c}{i5.CTM}\\
	& \multicolumn{1}{c}{\si{\mebi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} \\
	\midrule Base P4 & 6776.67 & 88.24 & 268.52 & 2.91 & 858.28 & 10.48 & 0 & 0 & 0 & 0 \\
	\Indfw(1) & 6778.77 & 88.27 & 1738.01 & 18.86 & 1263.28 & 15.42 & 23.50 & 36.72 & 90.00 & 35.16 \\
	\Indfw(4) & 6778.79 & 88.27 & 1746.88 & 18.95 & 1263.28 & 15.42 & 49.00 & 76.56 & 90.00 & 35.16 \\
	\Coopfw(1) & 6778.76 & 88.27 & 1742.51 & 18.91 & 1263.28 & 15.42 & 19.66 & 30.72 & 90.00 & 35.16 \\
	\Coopfw(4) & 6778.76 & 88.27 & 1738.04 & 18.86 & 1263.28 & 15.42 & 31.82 & 49.71 & 90.00 & 35.16 \\
	\bottomrule
\end{tabular}
}
\end{table}

@@

@@

\begin{appendices}
	\begin{figure}
		\resizebox{1.0\linewidth}{!}{
			\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
		}
		\caption{Latency! varying work.\label{fig:vary-work-latency}}
	\end{figure}

	\begin{figure}
		\resizebox{1.0\linewidth}{!}{
			\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
		}
		\caption{Latency! varying core count\label{fig:vary-core-latency}}
	\end{figure}
	\section{Old ``What could we do''?}
	How do we evaluate this without looking into classification performance?
	\begin{itemize}
		\item Analytically?
		\begin{enumerate}
			\item Prove that tile splitting scheme over memory hierarchy can improve performance.
			\begin{itemize}
				\item Hit rate of different tiles according to tile dimensionality?
				\item Access time for each tier of memory hierarchy.
			\end{itemize}
		\end{enumerate}
		\item Experimentally?
		\begin{enumerate}
			\item Create ``VNF'' testbed: mirror traffic from switch to a host for telemetry processing/flow state. Send flow state over to another node who computes RL actions. RL actions installed on switch.
			\begin{itemize}
				\item Why not co-host RL computation with telemetry? Could be separate functions, could want to ensure that neither impacts the performance of the other.
			\end{itemize}
			\item Create ``PDP'' testbed: Load firmware, pass in traffic. Traffic processing/action compute/table mod occurs in PDP.
			\item What traffic? Healthy mix of attack (UDP), Discord legit Opus VoIP traffic, HTTP bulk-transfer traffic.
			\item Measure $t_{1 \cdots 3}$ according to \cref{fig:state-slip} in both VNF and PDP contexts.
			\begin{description}
				\item[$t_1$]
				\begin{description}
					\item[PDP] Packet ingress $\rightarrow$ INT Island $\rightarrow$ State arrival in Control ME.
					\item[VNF] Packet mirror time $\rightarrow$ telemetry VNF $\rightarrow$ state arrival in agent vnf.
				\end{description}

>>>>

\newcommand{\approach}{On Path Learning}
\newcommand{\approachshort}{OPaL}
\newcommand{\Coopfw}{CoOp}
\newcommand{\coopfw}{\Coopfw}
\newcommand{\Indfw}{Ind}
\newcommand{\indfw}{\Indfw}

@@

\begin{figure}
	\centering
	\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/arch-with-p4}
	\caption{Architecture for an \approachshort{} agent on Netronome hardware, using 4 cores for parallel policy processing. (FIXME: redo in illustrator? Show CTXes, PISA, \Coopfw{} vs \Indfw{})\label{fig:netro-arch}}
\end{figure}

@@

\begin{figure*}
	\centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/ind}
		\caption{\Indfw{} (offline throughput-optimal).\label{fig:single-and-parallel:single}}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[keepaspectratio, width=0.9\linewidth]{figures/coop}
		\caption{\Coopfw{} (online-optimal).\label{fig:single-and-parallel:parallel}}
	\end{subfigure}
	\caption{\approachshort{}'s compute strategies scale to fit device capacity according to latency/throughput needs. ?? MAKE ME FOR REAL \label{fig:single-and-parallel}}
\end{figure*}

@@

\begin{figure}
	\resizebox{1.0\linewidth}{!}{
		\includegraphics{../plots/build/rl-perf-tester/vary-work-latency}
	}
	\caption{Latency! varying work.\label{fig:vary-work-latency}}
\end{figure}

\begin{figure}
	\resizebox{1.0\linewidth}{!}{
		\includegraphics{../plots/build/rl-perf-tester/vary-core-latency}
	}
	\caption{Latency! varying core count\label{fig:vary-core-latency}}
\end{figure}

\begin{figure}
	\centering
	\begin{subfigure}{0.45\linewidth}
		\resizebox{1.0\linewidth}{!}{
			\includegraphics{../plots/build/rl-perf-tester/latency-cumul}
		}
		\caption{Just Pars.}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\resizebox{1.0\linewidth}{!}{
			\includegraphics{../plots/build/rl-perf-tester/latency-cumul-broad}
		}
		\caption{Wider View.}
	\end{subfigure}
	\caption{BigCap}
\end{figure}

@@

\begin{table}
\caption{NFP memory use due to \approachshort{} using \numlist{1;4} MEs (\SI{32}{\bit}). CLS and CTM regions are shared between all programs on the same island (placing our RL agent on i5), while EMEM and IMEM regions are shared between all NFP programs on a NIC.\label{tab:resources} ??NEEDS REWORKED}
\resizebox{\linewidth}{!}{
\begin{tabular}{
		@{}c
		S[table-format=4.2] S[table-format=2.2]
		S[table-format=4.2] S[table-format=2.2]
		S[table-format=4.2] S[table-format=2.2]
		S[table-format=2.2] S[table-format=2.2]
		S[table-format=3.2] S[table-format=2.2]
		@{}
	}
	\toprule Firmware & \multicolumn{2}{c}{EMEM} & \multicolumn{2}{c}{EMEM Cache} & \multicolumn{2}{c}{IMEM} & \multicolumn{2}{c}{i5.CLS} & \multicolumn{2}{c}{i5.CTM}\\
	& \multicolumn{1}{c}{\si{\mebi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} & \multicolumn{1}{c}{\si{\kibi\byte}} & \multicolumn{1}{c}{\si{\percent}} \\
	\midrule Base P4 & 6776.67 & 88.24 & 268.52 & 2.91 & 858.28 & 10.48 & 0.00 & 0.00 & 0.00 & 0.00 \\
	\Indfw(1) & 6780.21 & 88.28 & 2541.07 & 27.57 & 1263.28 & 15.42 & 24.75 & 38.67 & 94.25 & 36.82 \\
	\Indfw(4) & 6780.21 & 88.28 & 2545.32 & 27.62 & 1263.28 & 15.42 & 51.18 & 79.97 & 107.00 & 41.80 \\
	\Coopfw(1) & 6779.12 & 88.27 & 1769.83 & 19.20 & 1263.28 & 15.42 & 20.25 & 31.64 & 90.00 & 35.16 \\
	\Coopfw(4) & 6779.12 & 88.27 & 1769.84 & 19.20 & 1263.28 & 15.42 & 31.91 & 49.85 & 90.00 & 35.16 \\
	\bottomrule
\end{tabular}
}
\end{table}

@@

?? Simialr HW has been used to get NN \emph{inference} \textcite{langlet-ml-netronome}---although \emph{actual NN latency} in his Fig 4.4 (IPG is inter-packet gap: his is not truly async but achieves perf by lying on the packet path. Note that 0 IPG murders his latencies).

@@

\begin{appendices}
	\section{Old ``What could we do''?}
	How do we evaluate this without looking into classification performance?
	\begin{itemize}
		\item Analytically?
		\begin{enumerate}
			\item Prove that tile splitting scheme over memory hierarchy can improve performance.
			\begin{itemize}
				\item Hit rate of different tiles according to tile dimensionality?
				\item Access time for each tier of memory hierarchy.
			\end{itemize}
		\end{enumerate}
		\item Experimentally?
		\begin{enumerate}
			\item Create ``VNF'' testbed: mirror traffic from switch to a host for telemetry processing/flow state. Send flow state over to another node who computes RL actions. RL actions installed on switch.
			\begin{itemize}
				\item Why not co-host RL computation with telemetry? Could be separate functions, could want to ensure that neither impacts the performance of the other.
			\end{itemize}
			\item Create ``PDP'' testbed: Load firmware, pass in traffic. Traffic processing/action compute/table mod occurs in PDP.
			\item What traffic? Healthy mix of attack (UDP), Discord legit Opus VoIP traffic, HTTP bulk-transfer traffic.
			\item Measure $t_{1 \cdots 3}$ according to \cref{fig:state-slip} in both VNF and PDP contexts.
			\begin{description}
				\item[$t_1$]
				\begin{description}
					\item[PDP] Packet ingress $\rightarrow$ INT Island $\rightarrow$ State arrival in Control ME.
					\item[VNF] Packet mirror time $\rightarrow$ telemetry VNF $\rightarrow$ state arrival in agent vnf.
				\end{description}
